Calendar-based graphics for visualizing people's daily schedules

### 2019-10-03 reviews/response.Rmd

> p11, l5: "The date argument specifies the date variable used to construct the calendar layout." Not very informative. Either omit or provide an example, e.g., date in which format (numerical, text, ...), from-to dates, etc.

TODO

> Figs 6-9 captions: You speak of "line glyphs", "line graphs", and "line charts". These mean the same - so be consistent and use just one of the three terms throughout the figure captions (and main text).

Done. We used "line graphs" consistently throughout the manuscript.

@@

> p22, l23: You wrote "boxes for household 2 indicate greater variability in daily energy consumption with noticeable variations on Thursdays," -- isn't this on Fridays?

TODO

> p22 vs. Fig 14: You speak of 6am, 6pm, 3pm, etc. in the text, but use 0, 6, 12, 18, 24 hour labels in the graph. Be consistent and use the international 0-24 hour notation in the text as well.

> Fig 15: It appears that April 30 is wrapped to the start of the monthly display? If so, add a comment similar to the one from Fig 6 to the figure caption.

### 2019-10-03 calendar-vis.Rmd
addressed many comments from reviewer 3

The proposed algorithm goes beyond the calendar-based heatmap. The approach is developed with three conditions in mind: (1) to display time-of-day variation in addition to longer temporal components such as day-of-week and day-of-year; (2) to incorporate line graphs and other types of glyphs into the graphical toolkit for the calendar layout; (3) to accentuate unusual patterns, such as those related to special events, for viewers. The proposed algorithm has been implemented in the `frame_calendar()` and `facet_calendar()` functions in the **sugrrants** package using R.

@@

* $k = 1, \dots , 7$ is the day of the week, that is the first day of the month.
* $d = 28, 29, 30$ or $31$ representing the number of days in any month.
* $(i, j)$ is the grid position where $1 \le i \le 5$, is week within the month, $1 \le j \le 7$, is day of the week.
* $g = k, \dots,(k+d)$ indexes the day in the month, inside the 35 possible cells.

@@

To create the layout for a full year, $(m, n)$ denotes the position of the month arranged in the plot, where $1 \le m \le M$ and $1 \le n \le N$; $b$ denotes the small amount of white space between each month for visual separation. Figure \@ref(fig:year-diagram) illustrates this layout where $M = 3$ and $N = 4$.

@@

(ref:fs-2016-cap) The calendar plot of hourly foot traffic at Flagstaff Station, using line glyphs. The disparities between weekday and weekend along with public holiday, are immediately apparent. The arrangement of the data into a $3 \times 4$ monthly grid represents all the traffic in 2016. Note that the algorithm wraps the last few days in the sixth week to the top row of each month block for a compact layout, which occurs in May and October.

Figure \@ref(fig:fs-2016) shows the line glyphs framed in the monthly calendar over the year 2016. This is achieved by the `frame_calendar()` function, which computes the coordinates on the calendar for the input data variables. These can then be plotted using the usual **ggplot2** R package [@R-ggplot2]. Thus, the grammar of graphics can be applied.

@@

(ref:fs-free-cap) Line glyphs on the calendar format showing hourly foot traffic at Flagstaff Station, scaled individually by day. The shape on a single day becomes more distinctive, as compared to Figure \@ref(fig:fs-2016).

@@

(ref:facet-cap) Line charts, embedded in the $6 \times 2$ monthly calendar, colored and faceted by the 3 sensors. The variations of an individual sensor are emphasized.

@@

Figure \@ref(fig:hod) shows energy consumption against time of day, separately by weekday and weekend. Household 1 is an early riser, starting their day before 6am and going back home around 6pm on weekdays. They switch air conditioning on when they get home from work and keep it operating until midnight, evident from the small horizontal cluster of points around 0.8 kWh. On the other hand, the stripes above 1 kWh for household 2 indicates that air conditioning may run continuously for some periods, consuming twice the energy as household 1. A third peak occurs around 3pm for household 3 only, likely coinciding when the children arrive home from school. They also have a consistent energy pattern between weekdays and weekends. As for household 4, their home routine starts after 6pm on weekdays. Figures \@ref(fig:dow) and \@ref(fig:hod), part of a traditional graphical toolkit, are useful for summarizing overall deviations across days and households.

@@

(ref:hod-cap) Scatterplots of half-hourly energy usage by time of day, with hourly averages overlaid, faceted by household and type of day. All households have different weekdays versus weekends daily routines. On weekdays, household 1 wakes up early before 6am, and household 2 around 6am, followed by household 3 and 4. The use of air conditioning is notable in households 1 and 2, as seen by horizontal clusters of points.

@@

Figure \@ref(fig:calendar-elec) displays the global scaling of each household's data in a calendar layout, unfolding their day-to-day life via electricity usage. Glancing over household 1, their overall energy use is relatively low. Their weekday energy use is distinguishable from their weekends, indicating a working household. The air conditioner appears to be used in the summer months (January and February) for a couple of hours in the evening and weekends. In contrast, household 2 keeps a cooling system functioning for much longer hours, which becomes more evident from late Wednesday through Thursday to early Friday in mid-January. These observations help to explain the stripes and clusters of household 2 in Figure \@ref(fig:hod). It is difficult to give a succinct description of household 3 since everyday energy pattern is variable, but May and June see more structure than the previous months. Individual data can be idiosyncratic, hence aggregated plots like Figure \@ref(fig:dow) and \@ref(fig:hod) are essential for assembling pieces to form a picture. However, the calendar plots tell the stories that are untold by previous plots, for example, their vacation time. Household 1 is on vacation over three weeks of mid-June, and household 2 also took some days off in the second week of June. Further, household 3 takes one short trip in January and the another one starting in the fourth week of June. Household 4 is away over two or three weeks in early April and late June. They all tend to take breaks during June probably due to the fact that the University winter break starts in June.

>>>>

The proposed algorithm goes beyond the calendar-based heatmap. The approach is developed with three conditions in mind: (1) to display time-of-day variation in addition to longer temporal components such as day-of-week and day-of-year; (2) to incorporate lines and other types of glyphs into the graphical toolkit for the calendar layout; (3) to accentuate unusual patterns, such as those related to special events, for viewers. The proposed algorithm has been implemented in the `frame_calendar()` and `facet_calendar()` functions in the **sugrrants** package using R.

@@

* $k = 1, \dots , 7$ is the day of the week, that is the first day of the month.
* $d = 28, 29, 30$ or $31$ representing the number of days in any month.
* $(i, j)$ is the grid position where $1 \le i \le 5$, is the row (week within the month), $1 \le j \le 7$, is the column (day of the week), with $(1, 1)$ being in the upper left corner.
* $g = k, \dots,(k+d)$ indexes the day in the month, inside the 35 possible cells.

@@

To create the layout for a full year, $(m, n)$ denotes the position of the month arranged in the plot, where $1 \le m \le M$ is the row and $1 \le n \le N$ is the column; $b$ denotes the small amount of white space between each month for visual separation. Figure \@ref(fig:year-diagram) illustrates this layout where $M = 3$ and $N = 4$.

@@

(ref:fs-2016-cap) The calendar plot of hourly foot traffic at Flagstaff Station, using line graphs. The disparities between weekday and weekend along with public holiday, are immediately apparent. The arrangement of the data into a $3 \times 4$ monthly grid represents all the traffic in 2016. Note that the algorithm wraps the last few days in the sixth week to the top row of each month block for a compact layout, which occurs in May and October.

Figure \@ref(fig:fs-2016) shows the line graphs framed in the monthly calendar over the year 2016. This is achieved by the `frame_calendar()` function, which computes the coordinates on the calendar for the input data variables. These can then be plotted using the usual **ggplot2** R package [@R-ggplot2]. Thus, the grammar of graphics can be applied.

@@

(ref:fs-free-cap) Line graphs on the calendar format showing hourly foot traffic at Flagstaff Station, scaled individually by day. The shape on a single day becomes more distinctive, as compared to Figure \@ref(fig:fs-2016).

@@

(ref:facet-cap) Line graphs, embedded in the $6 \times 2$ monthly calendar, colored and faceted by the 3 sensors. The variations of an individual sensor are emphasized.

@@

Figure \@ref(fig:hod) shows energy consumption against time of day, separately by weekday and weekend. Household 1 is an early riser, starting their day before 6 and going back home around 18 on weekdays. They switch air conditioning on when they get home from work and keep it operating until midnight, evident from the small horizontal cluster of points around 0.8 kWh. On the other hand, the stripes above 1 kWh for household 2 indicates that air conditioning may run continuously for some periods, consuming twice the energy as household 1. A third peak occurs around 15 for household 3 only, likely coinciding when the children arrive home from school. They also have a consistent energy pattern between weekdays and weekends. As for household 4, their home routine starts after 18 on weekdays. Figures \@ref(fig:dow) and \@ref(fig:hod), part of a traditional graphical toolkit, are useful for summarizing overall deviations across days and households.

@@

(ref:hod-cap) Scatterplots of half-hourly energy usage by time of day, with hourly averages overlaid, faceted by household and type of day. All households have different weekdays versus weekends daily routines. On weekdays, household 1 wakes up early before 6, and household 2 around 6, followed by household 3 and 4. The use of air conditioning is notable in households 1 and 2, as seen by horizontal clusters of points.

@@

Figure \@ref(fig:calendar-elec) displays the global scaling of each household's data in a calendar layout, unfolding their day-to-day life via electricity usage. Glancing over household 1, their overall energy use is relatively low. Their weekday energy use is distinguishable from their weekends, indicating a working household. The air conditioner appears to be used in the summer months (January and February) for a couple of hours in the evening and weekends. In contrast, household 2 keeps a cooling system functioning for much longer hours, which becomes more evident from late Wednesday through Thursday to early Friday in mid-January. These observations help to explain the stripes and clusters of household 2 in Figure \@ref(fig:hod). It is difficult to give a succinct description of household 3 since everyday energy pattern is variable, but May and June see more structure than the previous months. Individual data can be idiosyncratic, hence aggregated plots like Figure \@ref(fig:dow) and \@ref(fig:hod) are essential for assembling pieces to form a picture. However, the calendar plots tell the stories that are untold by previous plots, for example, their vacation time. Household 1 is on vacation over three weeks of mid-June, and household 2 also took some days off in the second week of June. Further, household 3 takes one short trip in January and another one starting in the fourth week of June. Household 4 is away over two or three weeks in early April and late June. They all tend to take breaks during June probably due to the fact that the University winter break starts in June.
